---
title: "Homework 4"
author: "Natalie Boettcher"
date: "2025-02-18"
output: 
  html_document:
    toc: true
    toc_float: true
    css: style.css
  
---

[Github Link](https://github.com/nboettcher111/SDS315)

```{r global options, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
```


```{r}

# load libraries
library(mosaic)
library(tidyverse)

```

## **Problem 1: Iron Bank**

**Null Hypothesis:** Securities trades from the Iron Bank are flagged 2.4% of the time, the same rate as other traders:

<div style="text-align: center;">
  Ho: P(flagged) = 0.024
</div>

**Test Statistic:** The test statistic is the number of flagged trades.

**Plot:** The following histogram shows the distribution of flagged trades under the null hypothesis:

```{r}

# 100000 simulations, 2021 trades, 0.024 probability flagged
simulatedTrades <- do(100000)*rflip(2021, prob = 0.024)
```

```{r}

# plot distribution of flagged trades
ggplot(simulatedTrades, aes(x = heads)) +
  geom_histogram(binwidth = 1, fill = "orchid3", color = "black", boundary = 0.5) +
  labs(title = "Distribution of Flagged Trades",
       x = "Number of Flagged Trades",
       y = "Frequency") +
  theme_minimal()
  
```

**P-Value:** The p-value is 0.0216, the proportion of simulations where the number of flagged trades is greater than or equal to 70.

```{r, results='hide'}

pValue <- prop(simulatedTrades$heads >= 70)
pValue

```

**Conclusion:** Since the p-value is less than 0.05, it is considered statistically significant. This suggests that 70 flagged trades is unlikely to have happened by chance if the null hypothesis is true.



## **Problem 2: Health Inspections**

**Null Hypothesis:** Health inspections at Gourmet Bites result in health code violations at a 3% rate:

<div style="text-align: center;"> Ho: P(violation) = 0.03 </div>

**Test Statistic:** The test statistic is the number of health code violations.

**Plot:** The following histogram shows the distribution of health code violations under the null hypothesis, assuming a 3% rate of violations:

```{r}

# 100000 simulations, 50 inspections, 0.03 probability violation
simulatedViolations <- do(100000) * rflip(50, prob = 0.03)

```

```{r}

# plot distribution of health code violations
ggplot(simulatedViolations, aes(x = heads)) +
  geom_histogram(binwidth = 1, fill = "seagreen4", color = "black", boundary = 0.5) +
  labs(title = "Distribution of Health Code Violations",
       x = "Number of Health Code Violations",
       y = "Frequency") +
  theme_minimal()


```

**P-Value:** The p-value is 0.00017, the proportion of simulations where the number of health code violations is greater than or equal to 8.

```{r, results='hide'}

pValueHealth <- prop(simulatedViolations$heads >= 8)
pValueHealth

```

**Conclusion:** Since the p-value is less than 0.05, it is statistically significant. This suggests that 8 violations is unlikely to have happened by chance if the null hypothesis is true. This could suggest that Gourmet Bites might have a higher rate of health code violations than the citywide average.



## **Problem 3: Evaluating Jury Selection for Bias**

**Null Hypothesis:** The jury group distribution is the same as the population distribution.

**Alternative Hypothesis:** The jury group distribution is different from the population distribution.

**Test Statistic:** A chi-square test is used to compare the observed jury distribution with the expected distribution based on the county’s population proportions.

```{r}

# expected and observed counts
expectedDistribution <- c(Group1 = 0.3, Group2 = 0.25, Group3 = 0.2, Group4 = 0.15, Group5 = 0.1)
observedCounts <- c(Group1 = 85, Group2 = 56, Group3 = 59, Group4 = 27, Group5 = 13)

# total num. of jurors
totalJurors <- sum(observedCounts)

# compare observed and expected counts
tibble(observed = observedCounts, expected = expectedDistribution * totalJurors)
```

**Plot:** The following histogram shows the distribution of the chi-square statistic under the null hypothesis:

```{r}

# calculate chi squared statistic
chi_squared_statistic = function(observed, expected) {
  sum((observed - expected)^2 / expected)
}

# Simulations
num_simulations = 10000
chi2_sim = do(num_simulations) * {
  simulated_counts = rmultinom(1, totalJurors, expectedDistribution)
  this_chi2 = chi_squared_statistic(simulated_counts, totalJurors * expectedDistribution)
  c(chi2 = this_chi2)
}

# Plot distribution
ggplot(chi2_sim) + 
  geom_histogram(aes(x = chi2), binwidth = 0.5, fill = "tomato3", color = "black") +
  labs(title = "Distribution of Chi-Squared Statistics Under Null Hypothesis",
       x = "Chi-Squared Statistic",
       y = "Frequency") +
  theme_minimal()


```

**Observed Chi-Squared Statistic:** The observed chi-squared statistic is 12.426.

```{r, results='hide'}

my_chi2 = chi_squared_statistic(observedCounts, totalJurors * expectedDistribution)
my_chi2


```

**P-Value:** The p-value is 0.0129.

```{r, results='hide'}

# Calculate p-value
p_value = prop(chi2_sim$chi2 >= my_chi2)
p_value


```

**Conclusion:** Since the p-value is less than 0.05, the null hypothesis is rejected. This suggests that the jury selection process is significantly different from the county's population distribution.


## **Problem 4: LLM Watermarking**


## Part A: the null or reference distribution

```{r}

# 1. Read the sentences
sentences <- readLines("brown_sentences.txt")

# 2. Preprocess the text

preprocess_text <- function(sentence) {
  sentence <- gsub("[^A-Za-z]", "", sentence)
  sentence <- toupper(sentence)
  
  # Count the occurrences of each letter (A-Z)
  letter_counts <- table(strsplit(sentence, NULL))
  
  
  all_letters <- letters[1:26]
  letter_counts <- as.numeric(letter_counts[all_letters])
  names(letter_counts) <- all_letters
  return(letter_counts)
}


letter_counts_list <- lapply(sentences, preprocess_text)

```

```{r}

# 4. Compare with expected count

expected_frequency <- c(A = 0.078688, B = 0.01621, C = 0.023583, D = 0.041729, E = 0.121905,
                        F = 0.02157, G = 0.019764, H = 0.062936, I = 0.068048, J = 0.001504,
                        K = 0.00863, L = 0.043464, M = 0.027408, N = 0.066694, O = 0.078634,
                        P = 0.016037, Q = 0.001046, R = 0.059903, S = 0.06438, T = 0.08826,
                        U = 0.031281, V = 0.009825, W = 0.0237, X = 0.001395, Y = 0.022843, Z = 0.000502)


```

```{r}
# 5. Compute chi-squared statistic

chi_squared_statistic <- function(observed, expected, sentence_length) {
  expected_counts <- expected * sentence_length
  sum((observed - expected_counts)^2 / expected_counts)
}


```

```{r}
# 6. Compile the distribution

chi_squared_stats <- numeric(length(sentences))

# Calculate the chi-squared statistic for each sentence
for (i in 1:length(sentences)) {
  observed_counts <- letter_counts_list[[i]]
  sentence_length <- sum(observed_counts)  
  
  chi_squared_stats[i] <- chi_squared_statistic(observed_counts, expected_frequency, sentence_length)
}

```


```{r}
# Create the null distribution
null_distribution <- chi_squared_stats
```


## Part B: checking for a watermark

```{r}
# 1. input sentences into r vector
sentences_to_check <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

# 2. reprocess sentences
letter_counts_check_list <- lapply(sentences_to_check, preprocess_text)

# 3. calculate chi squared for each sentnece
chi_squared_check_stats <- numeric(length(sentences_to_check))

for (i in 1:length(sentences_to_check)) {
  observed_counts <- letter_counts_check_list[[i]]
  sentence_length <- sum(observed_counts)  # Length of the sentence (total number of letters)
  
  chi_squared_check_stats[i] <- chi_squared_statistic(observed_counts, expected_frequency, sentence_length)
}

# 4. p value for each sentence

p_values <- sapply(chi_squared_check_stats, function(stat) mean(null_distribution >= stat))

# 5.  table of p-values
p_values_table <- data.frame(
  Sentence = 1:10,
  P_Value = round(p_values, 3)
)

print(p_values_table)




```













