---
title: "Homework 8"
author: "Natalie Boettcher"
date: "2025-04-14"
output: 
  html_document:
    toc: true
    toc_float: true
    css: style.css
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
```


[Github Link](https://github.com/nboettcher111/SDS315)


```{r}

# load necessary libraries
library(ggplot2)
library(dplyr)
library(boot)

# problem 1 dataset
creatinine <- read.csv("creatinine.csv")

# problem 2 dataset
covid <- read.csv("covid.csv")

# problem 3 dataset
milk <- read.csv("milk.csv")

```


## **Problem 1: Regression Warm-Up**


## A: What creatinine clearance rate should we expect for a 55-year-old? 


```{r}

# Fit linear regression model
model <- lm(creatclear ~ age, data = creatinine)
#summary(model)

```
Estimated regression equation: **y = 147.813 − 0.620 * age**

&nbsp;&nbsp;&nbsp;&nbsp;Now we just plug 55 into the equation:

  y = 147.813 − 0.620 * 55 = 147.813 − 34.09 = 113.723
  
&nbsp;&nbsp;&nbsp;&nbsp;The expected creatinine clearance rate for a 55-year-old is about **113.72 mL/min.**


## B: How does creatinine clearance rate change with age?


This is simply the slope of the model. From part A, we know that the slope is about **−0.620**, Which means that the creatinine clearance rate decreases by 0.620 mL/min per year of age.


## C: Whose creatinine clearance rate is healthier (higher) for their age: a 40-year-old with a rate of 135, or a 60-year-old with a rate of 112? 


We need to Calculate expected creatinine clearance for each age using the equation: y = 147.813 − 0.61982 * age
  
  
&nbsp;&nbsp;&nbsp;&nbsp;For the **40-year-old:**

<div align="center">
y(40) = 147.813 − 0.61982 * 40 = 147.813 − 24.7928 = **123.0202**
</div>
  
&nbsp;&nbsp;&nbsp;&nbsp;Subtract from actual:

<div align="center"> 
135 - 123.02 = **11.98**
</div>  
  
  
&nbsp;&nbsp;&nbsp;&nbsp;For the 60-year-old: 

<div align="center">
y(60) = 147.813 − 0.61982 * 60 = 147.813 − 37.1892 = **110.6238**
</div> 
  
&nbsp;&nbsp;&nbsp;&nbsp;Subract from actual:

<div align="center">
112 - 110.62 = **1.38**
</div>
  
  
**Conclusion:** The 40-year-old has a clearance 11.98 mL/min above expected for their age, while the 60-year-old is only 1.38 mL/min above expected. Therefore, the 40-year-old has a healthier creatinine clearance rate for their age.<br><br>



## **Problem 2: Modeling Disease Growth**


## Model For Italy

```{r}

# model for Italy
italy <- covid %>% filter(country == "Italy", deaths > 0)
italyModel <- lm(log(deaths) ~ days_since_first_death, data = italy)
#summary(italyModel)
```
```{r}

# italy model growth rate
italyGrowth <- coef(italyModel)[2]
italyDoubling <- log(2) / italyGrowth

```

```{r}

# Italy bootstrap confidence interval
italyBoot <- function(data, indices) {
  d <- data[indices, ]
  fit <- lm(log(deaths) ~ days_since_first_death, data = d)
  growth <- coef(fit)[2]
  doubling <- log(2) / growth
  return(c(growth, doubling))
}

italyResults <- boot(italy, statistic = italyBoot, R = 1000)

# CI
#growth rate
#boot.ci(italyResults, index = 1, type = "perc")
#doubling time
#boot.ci(italyResults, index = 2, type = "perc")  

```


The exponential growth model for Italy yielded an estimated growth rate of **0.183** (95% CI: 0.1588,  0.2079), corresponding to a doubling time of **3.78** days (95% CI: 3.334,  4.364).

## Model For Spain

```{r}
# model for Spain
spain <- covid %>% filter(country == "Spain", deaths > 0)
spainModel <- lm(log(deaths) ~ days_since_first_death, data = spain)
#summary(spainModel)

```

```{r}

# Spain model growth rate
spainGrowth <- coef(spainModel)[2]
spainDoubling <- log(2) / spainGrowth

```

```{r}

# Spain bootstrap confidence interval
spainBoot <- function(data, indices) {
  d <- data[indices, ]
  fit <- lm(log(deaths) ~ days_since_first_death, data = d)
  growth <- coef(fit)[2]
  doubling <- log(2) / growth
  return(c(growth, doubling))
}


SpainResults <- boot(spain, statistic = spainBoot, R = 1000)

# CI
#boot.ci(SpainResults, index = 1, type = "perc")  # growth rate
#boot.ci(SpainResults, index = 2, type = "perc")  # doubling time
 

```

For Spain, the estimated growth rate was **0.276** (95% CI: 0.2335, 0.3172), with a doubling time of **2.51 days** (95% CI: 2.185,  2.968).


The following line plot shows daily reported deaths over time, showing the exponential growth during the first wave of the COVID-19 pandemic:


```{r}

# plot of daily deaths

ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1) +
  labs(
    title = "Daily COVID-19 Deaths in Italy and Spain",
    x = "Days Since First Death",
    y = "Daily Deaths"
  ) +
  theme_minimal()

```

## **Problem 3: Price Elasticity of Demand**

```{r}

# regression model
model <- lm(log(sales) ~ log(price), data = milk)
#summary(model)

```

```{r}

# Bootstrap function
elasticityBoot <- function(data, indices) {
  d <- data[indices, ]
  fit <- lm(log(sales) ~ log(price), data = d)
  return(coef(fit)[2])  
}


bootResults <- boot(milk, statistic = elasticityBoot, R = 1000)

# 95% CI
#boot.ci(bootResults, type = "perc")

```

To estimate the price elasticity of demand for milk, a power-law model was used of the form Q = KP^B,
which becomes linear in logs: 

<div align="center">
log(Q) = log(K) + B*log(P)
</div>

This model was estimated using linear regression, with log(sales) as the outcome and log(price) as the predictor.

The estimated price elasticity of demand is approximately **β = -1.620**, with a 95% bootstrapped confidence interval of **(-1.787, -1.460).**

This indicates that demand for milk is **elastic**, since the elasticity is less than -1. Consumers respond strongly to price changes: a 1% increase in milk price leads to approximately a 1.62% decrease in quantity demanded.



